The role of cilia-mediated cerebrospinal fluid flow in intraventricular transport in zebrafish embryos

### 2025-02-03 paper/main.tex
Edits in Results 1

\mer{@HH: I suggest you write \cref{fig:figure1}\textbf{a}, rather than \cref{fig:figure1} panel \textbf{a}, throughout. Increases signal-to-noise ratio.}

@@

\subsection*{Computational flow scenarios}
%\label{subsec:model_versions}
To study the impact of the motile cilia and the cardiac pulsations on the CSF flow, we consider the following computational flow scenarios, each predicting a CSF flow velocity $\uu$ and CSF pressure $p$.
\begin{itemize}
    \item A \emph{baseline} flow model representing a wild-type zebrafish including contributions both from the cilia forces and the cardiac pulsations. In this scenario, we apply the tangential traction $\btau$ on the cilia boundary $\Gc$ and prescribe the normal pressure $\Tilde{p}$ on the anterior/posterior boundary $\Gp$. % Model 0

@@

\subsection*{Cerebrospinal fluid flow in the brain ventricles is compartmentalized}

\mer{Why?} To characterize CSF flow patterns predicted by
computational fluid dynamics within the brain ventricles, we consider
the simulated CSF velocity and pressure when driven by both cilia and
cardiac pulsations. \mer{MER: Stopped here...} With cilia forces and
cardiac forces represented in the flow model (model version 0), we
observe a vortex-structured flow pattern in the anterior and middle
ventricles~(\cref{fig:figure1}, panel \textbf{e}). This rotational
motion shows that cilia-mediated flow is dominating in these regions,
as the vortex structures are a consequence of the applied tangential
stresses on the ciliated boundaries combined with the assumption of
fluid incompressibility. In the posterior ventricle, we observe
directional flow as the pressure-driven flow is dominating in this
region. The cardiac pressure forces induce a pulsatile motion of the
cerebrospinal fluid synchronized with the cardiac cycle, with velocity
magnitudes of the pulsatile motion being approximately an order of
magnitude smaller than the velocities related to the cilia-driven
motion. Over a cardiac cycle, the maximum velocity magnitude in the
ventricles was 27.7 $\mu$m/s, located dorsally in the posterior part
of the middle ventricle. This maximum velocity was attained at the
time instant coinciding with the peak of the cardiac pulsatile flow
flowing caudally, as a result of both the tangential stress boundary
condition and the pulsatile pressure-driven flow being aligned with
the rostrocaudal axis (the $x$ axis). A calculation using the observed
maximum velocity of 27.7 $\mu$m/s, and choosing the approximate height
of the middle ventricle $110 \ \mu$m as a characteristic length scale,
yields a Reynolds number $\mathrm{Re} \approx 0.004$, justifying the
assumption of Stokes flow.

\subsection*{Vortex structures vanish in the absence of cilia}
The pulsatile motion is absent when disregarding cardiac motion~(model I, \cref{fig:figure1}, panel~\textbf{f}). For model I, the boundary conditions are not time-dependent, rendering the solution to the Stokes equations~\eqref{eq:stokes_eqs} a steady-state flow field. Streamlines demonstrate the compartmentalization of the cerebrospinal fluid flow driven by cilia, and we observe highest vorticity in the vicinity of cilia. Excluding the cilia tangential traction from the model (model II), vortex-structured compartmentalization of the ventricular flow does not occur, and the flow is directional and pulsatile as a result of the pressure boundary conditions imposed~(\cref{fig:figure1}, panel \textbf{g}).

>>>>

\mer{@HH: I suggest you write \cref{fig:figure1}a, rather than \cref{fig:figure1} panel \textbf{a}, throughout. Increases signal-to-noise ratio.}

@@

\subsection*{Computational modelling scenarios}
%\label{subsec:model_versions}
To study the impact of the motile cilia and the cardiac pulsations on the CSF flow, we consider the following computational flow scenarios, each predicting a CSF flow velocity $\uu$ and CSF pressure $p$.
\begin{itemize}
    \item A \emph{baseline} flow model representing a wild-type zebrafish including contributions both from the cilia forces and the cardiac pulsations. In this scenario, we apply the tangential traction $\btau$ on the cilia boundary $\Gc$ and prescribe the normal pressure $\Tilde{p}$ on the anterior/posterior boundary $\Gp$. % Model 0

@@

\subsection*{Cilia induces flow compartmentalization within the brain ventricles}

\mer{One sentence on why? - what do we not know - what do we try to
  find out?} To characterize CSF flow patterns predicted by
computational fluid dynamics within the brain ventricles, we begin by
considering the simulated CSF velocity and pressure when driven by
both cilia motion and cardiac pressure pulsations. These forces induce
\fixme{persistent} flow features, with vortex-structures in the
anterior and middle ventricles and \fixme{unidirectional} flow in the
posterior ventricle (\Cref{fig:figure1}\textbf{e},
\fixme{Supplementary video 1}). These characteristic patterns suggest
that the flow in the anterior and middle ventricles are dominated by
the cilia contribution, while the cardiac pressure pulsations dominate
within the posterior ventricle. Over one cardiac cycle, the CSF flow
speed reaches up to 27.7 \textmu m/s, dorsally in the posterior part
of the middle ventricle. This peak is attained at the time instant
coinciding with the time of peak caudal cardiac pulsatile flow, as a
result of both the traction induced by the cilia and the pulsatile
pressure-driven flow being aligned with the rostrocaudal ($x-$)
axis. Using this maximum velocity magnitude and choosing the height of
the middle ventricle ($110$ \textmu m) as a characteristic length
scale, yields a Reynolds number $\mathrm{Re} \approx 0.004$,
justifying the assumption of Stokes flow. To further examine the
effect of the cilia, we next consider a no-cilia model variation. In
this scenario, the simulated CSF flow is pulsatile and
\fixme{unidirectional} (\cref{fig:figure1}g), with a peak velocity
magnitude of \fixme{$4.8$ \textmu m/s}. On the other hand, the
vortices (re)appear when considering a variation with cilia but no
cardiac contribution (\cref{fig:figure1}h), in which case the highest
velocities ($26.5$ \textmu m/s) also occur in the vicinity of the
cilia. Thus, the bidirectional traction induced by motile cilia
inherently leads to a partial compartmentalization of the ventricular
system.

%% %\subsection*{Vortex structures vanish in the absence of cilia}
%% The pulsatile motion is absent when disregarding cardiac motion~(model I, \cref{fig:figure1}, panel~\textbf{f}). For model I, the boundary conditions are not time-dependent, rendering the solution to the Stokes equations~\eqref{eq:stokes_eqs} a steady-state flow field. Streamlines demonstrate the compartmentalization of the cerebrospinal fluid flow driven by cilia, and we observe highest vorticity in the vicinity of cilia.

%% Excluding the cilia tangential traction from the model (model II), vortex-structured compartmentalization of the ventricular flow does not occur, and the flow is directional and pulsatile as a result of the pressure boundary conditions imposed~(\cref{fig:figure1}, panel \textbf{g}).

%% The cardiac pressure forces induce a pulsatile motion of the
%% cerebrospinal fluid synchronized with the cardiac cycle, with velocity
%% magnitudes of the pulsatile motion being approximately an order of
%% magnitude smaller than the velocities related to the cilia-driven
%% motion.
